\section{Conclusion}
\label{sec:conclusion}

StarNet employs forward variational inference and is more accurate than both a recently published MCMC-based probabilistic cataloger and a widely used non-model-based
procedure.
In the framework of probabilistic modeling,
StarNet produces catalog uncertainties captured by a posterior over the set of all catalogs.
Importantly, unlike current MCMC approaches, StarNet also has the capacity to scale probabilistic cataloging to process large astronomical surveys.

The quality of StarNet detections is the result of optimizing the forward KL, a different objective than the one traditionally used in variational inference.
Optimizing the forward KL allows the variational posterior to be fit on large amounts of complete data (i.e., images along with their latent catalogs) generated from StarNet's statistical model.
% While labeled data from simulators has been used in other astronomy applications to train deep neural networks (see for example~\cite{Lanusse_2017_cmudeeplens} and \cite{huang2019finding}), StarNet is the first training procedure to employ simulated data in a statistical framework: the neural network specifies an approximate Bayesian posterior.

While this work focuses on stars, our methodology can be extended to
include more general light sources, such as galaxies.
One promising direction is to incorporate a highly accurate deep generative model of galaxies~\citep{Regier2015ADG, Reiman_2019_gans_deblend, lanusse2020deep, Arcelin_2020} into the StarNet model.
% Here, a neural network encodes a conditional likelihood of galaxy images given a low-dimensional galaxy representation,
% and we estimate the the galaxy model by re-incorporating the ELBO objective. The wake-sleep algorithm is a method that alternates between optimizing
% the expected forward KL (``sleep" phase) to
% infer catalogs and optimizing the ELBO (``wake" phase) to fit the galaxy model \citep{Hinton1995wake_sleep}.
% Unsurprisingly, the current performance of StarNet is sub-optimal for cataloging
% regions of the
% sky that contain both stars and galaxies, due to model misfit (Appendix~\ref{sec:results_sparse_field}).
The statistical framework in this research lays the foundation for building flexible models to incorporate the cataloging of other celestial objects.

Future astronomical surveys will produce far more data than past surveys.
As telescopes peer deeper into space, fields will reveal more sources and images will become more crowded.
The uncertainties in crowded fields necessitate a probabilistic approach.
Our method holds the promise of providing  a scalable inference tool that can meet the challenges of future surveys.
